Semi-Permeable Vesicles: Mathematical Formulation, Numerics, and Physics

### 2020-08-17 vesicle_code/docs/permeableVes/main.tex
worked on pnas paper

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Dimensionless equations}
The inextensibility condition guarantees that the vesicle's length, $L$,
is conserved, but its area, $A$, is no longer conserved because of
semi-permeability. Therefore, the reduced area $\nu = 4\pi A/L^2 \in
(0,1]$ is not conserved. We non-dimensionalize~\eqref{eqn:vesVelocity}
by introducing the length scale $R_0 = L/2\pi$ and defining the time
scale $\tau = \mu R_0^3/k_b$, the tension scale $Q = k_b/R_0^2$, and the
permeability scale $W = R_0/\mu$. Since the bending stiffness $k_b$ only
sets the time scale, we let $k_b=1$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Area dynamics}
Following the analysis of Veerapaneni et al.~\cite{vee-raj-bir-pur2009},
we compute the steady-state shape of a semi-permeable vesicle in a
quiescent flow. The steady-state shape is a stationary point of the
Lagrangian
\begin{align}
  \mathcal{L} = \frac{1}{2}\int_{\gamma} \kappa^2 \, ds +
    \Lambda \left(\int_{\gamma} ds  - L \right),
\end{align}
where the Lagrange multiplier $\Lambda$ the constant tension. Note that
there is no pressure Lagrange multiplier since the vesicle's area
fluctuates. Taking the variation of $\mathcal{L}$ with respect to
$\gamma$, and setting this vesicle force to zero, the steady state shape
satisfies $\kappa_{ss} + \frac{1}{2}\kappa^3 - \Lambda \kappa = 0$,
with initial conditions $\kappa(0) = \kappa_0$ and $\kappa_s(0) = 0$.
Integrating once, we have
\begin{align}
  \frac{\kappa_s^2}{2} + \frac{\kappa^4}{8} - 
    \frac{\Lambda}{2}\kappa^2 = \kappa_0^4 - \frac{\Lambda}{2}\kappa_0^2.
  \label{eqn:curvODE}
\end{align}
Assuming $\kappa_s \neq 0$, this differential equation is separable, and
can be solved analytically in terms of the {\em EllipticF} function
using Mathematica. Unlike in the impermeable
case~\cite{vee-raj-bir-pur2009}, the solution of~\eqref{eqn:curvODE} can
not be made periodic unless $\kappa(s) = \kappa_0$. Therefore, the
steady state shape of a semi-permeable vesicle in a quiescent flow is a
circle. By the inextensibility condition, the steady state vesicle has a
radius of $L/2\pi$, so its curvature and tension are
\begin{align}
  \kappa = \frac{2\pi}{L}, \: \Lambda = \frac{2\pi^2}{L^2}.
  \label{eqn:SSshape}
\end{align}
Note that in contrast to a impermeable vesicle, the final tension is
positive rather than negative.

In addition to computing the steady shape, we can compute the transient
area of the vesicle. The area of the vesicle is
\begin{align}
  A(t) = \frac{1}{2}\int_{\gamma} (\xx \cdot \nn)\, ds,
  \label{eqn:area}
\end{align}
so the derivative of $A(t)$ includes terms due to the time derivative of
$\xx \cdot \nn$, and a term due to interface
stretching~\cite{lai-tse-hua2008}, but the stretching term is zero
because of the local inextensibility condition. Therefore,
\begin{align}
  \dot{A}(t) =
  \frac{1}{2} \int_{\gamma} (\dot{\xx} \cdot \nn)\, ds  + 
  \frac{1}{2} \int_{\gamma} (\xx \cdot \dot{\nn})\, ds.
\end{align}
Applying integration by parts to the second integral, it is equivalent
to the first.
%\begin{align}
%  \int_\gamma (\xx \cdot \dot{\nn})\, ds 
%    = \int_\gamma \left(\xx \cdot \dot{\xx}_s^\perp \right)
%    = -\int_\gamma \left(\xx^\perp \cdot \dot{\xx}_s\right)\, ds 
%    = \int_\gamma \left(\xx_s^\perp \cdot \dot{\xx}\right)\, ds 
%    = \int_\gamma (\dot{\xx} \cdot \nn)\, ds.
%\end{align}
Therefore,
\begin{align}
  \dot{A}(t) = \int_\gamma \left(\beta (\ff \cdot \nn)\nn 
    + \SS[\ff]\right) \cdot \nn\, ds 
  = \beta \int_\gamma (\ff \cdot \nn)\, ds,
\end{align}
where the last equality follows from the incompressibility of the
single-layer potential.  Using integration by parts and the
Frenet-Serret identities, we have
\begin{align}
  \dot{A}(t) &= \beta \int_\gamma \left( -\xx_{ssss} + 
    (\sigma \xx_s)_s \right) \cdot \nn \, ds \\
  &= -\beta \int_\gamma \left(-\xx_{sss} + \sigma \xx_s 
    \right) \cdot \kappa \xx_s \, ds \\
  &= -\beta \int_\gamma \left(\xx_{ss} \cdot 
    (\kappa \xx_s)_s + \kappa \sigma \right) \, ds \\
  &= -\beta \int_\gamma \left(\kappa (\xx_{ss} \cdot \xx_{ss}) + 
    \kappa_s (\xx_{ss} \cdot \xx_s) + \kappa \sigma \right) 
    \, ds \\
  &= -\beta \int_\gamma \left(\kappa^3 + \kappa \sigma \right) 
    \, ds \\
  &= \beta \int_\gamma \left(
    \frac{\kappa^3}{2} - \kappa \Lambda \right) \, ds.
  \label{eqn:areaROC}
\end{align}
Note that this is consistent with the steady state curvature and tension
in~\eqref{eqn:SSshape} since these constant values result in the
integrand in~\eqref{eqn:areaROC} being zero.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Discretization in Space}
By using an integral equation formulation, the unknowns are the vesicle
position and tension which we track at a set of $N$ discretization
points. All arclength derivatives are computed with Fourier
differentiation, and eighth-order quadrature is applied to the weakly
singular single-layer potential~\cite{alp1999}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Discretization in Time}
The high-order derivatives in the bending and tension
forces~\eqref{eqn:membraneForces} result in stiffness if an explicit
time stepping method is applied to equation~\eqref{eqn:vesVelocity}.
Therefore, we use a modification of a time stepping method that
discretizes high-order derivatives
semi-implicitly~\cite{vee-gue-zor-bir2009}. Introducing notation for
all the required differential and projection operators
\begin{alignat}{3}
  &B[\xx]\ff = -\frac{d^4}{ds^4} \ff,  \qquad
  &&T[\xx]\sigma = (\sigma \xx_s)_s, \\
  &D[\xx]\ff = \xx_s \cdot \ff_s, 
  &&P[\xx]\ff = (\ff \cdot \nn) \nn. 
\end{alignat}
the dynamics of the vesicle are governed by
\begin{align}
  \dot{\xx} = \uu_{\infty}(\xx) + 
  \beta P(B\xx + T\sigma) + \SS(B\xx + T\sigma), \quad
  D \dot{\xx} = 0.
\end{align}
Letting $B^N$ be the bending operator due to the vesicle
configuration at time $t^N$, and similar notation for the other
operators, we use the semi-implicit time stepping method
\begin{alignat}{3}  
  \frac{\xx^{N+1} - \xx^N}{\Delta t} &= \uu_\infty(\xx^N) 
  &&+ \beta P^N\left(B^N\xx^{N+1} + T^N\sigma^{N+1}\right) \nonumber \\
  & &&+ \SS^N\left(B^N\xx^{N+1} + T^N\sigma^{N+1}\right), \\
  D^N\xx^{N+1} &= 1.
\end{alignat}

To achieve high-order accuracy over long time horizons, we modify the
adaptive spectral deferred correction time stepping method introduced
in~\cite{qua-bir2016}. Since the area of the vesicle is no longer
conserved, we only use the length of the vesicle to estimate the
local truncation error which then determines if a time step is accepted
or rejected and the subsequent time step size.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Quiescent Flow} 
We consider a semi-permeable elliptical vesicle with initial reduced
area $\nu = 0.65$ and several different permeability rates in a
quiescent flow. The steady-state of an impermeable vesicle has constant
negative tension~\cite{kra-win-sei-lip1996}, but based
on~\eqref{eqn:SSshape}, a semi-permeable vesicle relaxes to a circle
with positive constant tension $\Lambda = 2\pi^2/L^2$. Recalling the
governing equation for the area in~\eqref{eqn:areaROC}, we expect both
the time for the vesicle to deviate from its initial area, and the time
to reach its steady state will scale with $\beta^{-1}$. In
Table~\ref{tbl:ellipseRelaxTimes}, we report the time,
$t_\mathrm{steady}$, required for the vesicle's area to be within 1.3\%
of its maximum area, and the time, $t_\mathrm{inflate}$, required for
the vesicle's area to exceed 1.3\% of its initial area. We see that both
times scale with $\beta^{-1}$. We also plot the area of a vesicle with
six different permeability rates in Figure~\ref{fig:ellipseArea}. The
area is computed directly from the vesicle shape (solid lines) and by
integrating equation~\eqref{eqn:areaROC} (marks) with a first-order
method. We also include black dashed lines indicating the times,
$t_{\mathrm{steady}}$, that the vesicle is near its steady state
circular shape.
\begin{table}[htp]
  \centering
  \caption{\label{tbl:ellipseRelaxTimes} The time for a semi-permeable
  vesicle to its steady state ($t_\mathrm{steady}$) and to begin
  inflating ($t_\mathrm{inflate}$).}
  \begin{tabular}{C{0.8cm}C{2cm}C{0.8cm}C{2cm}C{0.8cm}}
    $\beta$ & $t_\mathrm{steady}$ & ratio & $t_\mathrm{inflate}$ & ratio \\
    \midrule
    $10^{0}$  & $6.45 \times 10^{0}$ & ---  
              & $2.17 \times 10^{-2}$ & --- \\ 
    $10^{-1}$ & $2.00 \times 10^{1}$ & 3.10 & 
                $1.25 \times 10^{-1}$ & 5.76 \\
    $10^{-2}$ & $9.61 \times 10^{1}$ & 4.81 & 
                $1.59 \times 10^{0}$ & 12.72 \\
    $10^{-3}$ & $8.36 \times 10^{2}$ & 8.71 & 
                $2.19 \times 10^{1}$ & 13.77 \\
    $10^{-4}$ & $8.15 \times 10^{3}$ & 9.74 & 
                $2.14 \times 10^{2}$ & 9.77 \\
    $10^{-5}$ & $8.03 \times 10^{4}$ & 9.85 & 
                $2.20 \times 10^{3}$ & 10.14 \\
    \bottomrule
  \end{tabular}
\end{table}

\begin{figure}[htp]
  \centering
%  \ifTikz
%  \input{figures/ellipseArea.tikz}
%  \fi
  \includegraphics[width=0.8\linewidth]{figures/ellipseArea}
  \caption{\label{fig:ellipseArea} The area of an elliptical
  semi-permeable vesicle, with permeability rate $\beta$ in a quiescent
  flow. The solid lines compute the area using
  equation~\eqref{eqn:area}, while the marks are computed by integrating
  equation~\eqref{eqn:areaROC}. The dashed black lines are the times
  when the vesicle's area is within 1.3\% of its maximum value of
  $4\pi$.}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Shear Flow}
\begin{itemize}
  \item Mild dependence on initial shape

  \item Look at tank-treading angle, tank-trading speed, reduced area,
    absolute value of the permeable flux

  \item Max flux at the steady state as a function of beta and
    $\dot{\gamma}$

  \item reduced area/inclination angle for a particular $(\beta,\gamma)$
    and different flux rates using the gating functions (Apr 7 and 9) as
    well as constant $\beta$
\end{itemize}

We consider a semi-permeable vesicle in the shear flow
$\uu_{\infty}(\xx) = \dot{\gamma} (y,0)$ with different permeability
rates $\beta$ and shear rates $\dot{\gamma}$.  For a two-dimensional
impermeable vesicle, depending on the reduced area and viscosity
contrast, vesicles in a shear flow either tank tread or
tumble~\cite{fin-lam-sei-gom2008,kra-win-sei-lip1996}, but tumbling is
not possible with viscosity contrast of one. Since we are considering
semi-permeable vesicles, we assume that there is no viscosity contrast,
and we do not expect to observe tumbling dynamics.

We initialize an elliptical vesicle with reduced area $\nu = 0.65$, and
with its long axis oriented in the $y$-direction. For all permeability
rates and shear rates, the semi-permeable vesicle tilts to an
inclination angle and undergoes tank-treading behavior. However, because
of the semi-permeability, the final reduced area does not agree with the
initial reduced area. In Figure~\ref{fig:shearPhaseDiagram}, we plot the
final shape of the vesicle for various permeability rates, $\beta$, and
shear rates $\dot{\gamma}$. We also includes lines of constant final
reduced area. We see that the shear rate has the largest effect on the
final reduced area. However, for small $\beta$, the permeability rate
also has a large effect on the final reduced area. In particular, the
smallest final reduced areas are the result of large shear rates and
small permeability rates.
\todo[inline]{Also reporting inclination angle as well}

\begin{figure}[htp]
  \centering
  %\ifTikz
  %\input{figures/shearPhaseDiagramRA2.tikz}
  %\fi
  \includegraphics[width=0.9\linewidth]{figures/shearPhaseDiagramRA2.pdf}
  \caption{\label{fig:shearPhaseDiagram} The steady state shape of a
  semi-permeable vesicle with several different permeability rates
  $\beta$ and shear rates $\dot{\gamma}$. The black lines represent
  contour lines of the final reduced areas (left) and the inclination
  angle (right). The reduced area and inclination angle of each contour
  line are reported.}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Stenosis}
We consider a semi-permeable vesicle in a constricted geometry. A
double-layer potential is used to impose a boundary condition on the
outer wall, and near-singular integration~\cite{qua-bir2014} quadrature
resolves the vesicle-wall interaction.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Poiseuille Flow}
We consider a semi-permeable vesicle in the Poiseuille flow $\uu =
U(1-(y/R)^2,0)$, where $R=10$ determines the flow width, and $U$ is the
maximum flow rate. The length of the vesicle is $L = 4\pi$ and its
initial reduced area is $\nu = 0.65$. We first investigate the the
effect of the flow rate $U$ and the permability rate $\beta$ on the
steady state shape of the vesicle when it is initially placed on the
center line $y=0$. Figure~\ref{fig:parabolicPhaseDiagram} shows the
steady state configuration for a variety of flow rates $U$ and
permeability rates $\beta$. Similar to the shear example, we see that
the flow rate has the biggest effect on the final shape while the
permeability rate has a much weaker effect. Instead, the permeability
rate determines the time required for the vesicle to reach a steady
state. In Tables~\ref{tbl:parabolicRelaxTimes1}
and~\ref{tbl:parabolicRelaxTimes2}, we consider two flow rates and
report the time, $t_\mathrm{steady}$, required for the vesicle's area to
be within 1.3\% of its maximum area, and the time, $t_\mathrm{inflate}$,
required for the vesicle's area to exceed 1.3\% of its initial area. We
see that both times scale with $\beta^{-1}$.

As a second experiment, we investigate the effect of $U$ and $\beta$
when the vesicle is initialized above line $y=0$. Kaoui et
al.~\cite{kao-bir-mis2009} analyzed the migration of the same
experiment, but with an impermeable vesicle. They showed that for
above a critical reduced area, the vesicle migrates to the center line
$y=0$ as a symmetric shape. However, for smaller reduced areas, the
limiting shape is a tank-treading slipper with a center of mass above
the center line $y=0$. We repeat this experiment but with a
semi-permeable vesicle.

\begin{figure}[htp]
%  \ifTikz
%  \input{figures/parabolicPhaseDiagram.tikz}
%  \fi
  \includegraphics[width=0.9\linewidth]{figures/parabolicPhaseDiagram.pdf}
  \caption{\label{fig:parabolicPhaseDiagram} The steady state shape of a
  semi-permeable vesicle with several different permeability rates
  $\beta$ and flow rates $U$. At high flow rates, the vesicle
  self-intersects before reaching a steady state configuration.}
\end{figure}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Dimensionless equations}
The inextensibility condition guarantees that the vesicle's length, $L$,
is conserved, but its area, $A$, is not because of semi-permeability.
Therefore, the reduced area $\nu = 4\pi A/L^2 \in (0,1]$ is not
conserved. We non-dimensionalize~\eqref{eqn:vesVelocity} by introducing
the length scale $R_0 = L/2\pi$ and defining the time scale $\tau = \mu
R_0^3/k_b$, the tension scale $Q = k_b/R_0^2$, and the permeability
scale $W = R_0/\mu$. The background velocity $\uu_\infty$ is
non-dimensionalized with $R_0/\tau$. For all simulations, the bending
modulus is $10^{-19}J$. Using these scales, all equations are written in
dimensionless form.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Area dynamics}
Following the analysis of Veerapaneni et al.~\cite{vee-raj-bir-pur2009},
we compute the steady-state shape of a semi-permeable vesicle in a
quiescent flow. The steady-state shape is a stationary point of the
Lagrangian
\begin{align}
  \mathcal{L} = \frac{1}{2}\int_{\gamma} \kappa^2 \, ds +
    \Lambda \left(\int_{\gamma} ds  - L \right),
\end{align}
where the Lagrange multiplier $\Lambda$ is the tension. Note that there
is no pressure Lagrange multiplier since the vesicle's area fluctuates.
Taking the variation of $\mathcal{L}$ with respect to $\gamma$ and
setting it to zero, the steady state shape satisfies $\kappa_{ss} +
\frac{1}{2}\kappa^3 - \Lambda \kappa = 0$, with initial conditions
$\kappa(0) = \kappa_0$ and $\kappa_s(0) = 0$.  Integrating once, we have
\begin{align}
  \frac{\kappa_s^2}{2} + \frac{\kappa^4}{8} - 
    \frac{\Lambda}{2}\kappa^2 = \frac{\kappa_0^4}{8} - 
    \frac{\Lambda}{2}\kappa_0^2.
  \label{eqn:curvODE}
\end{align}
Assuming $\kappa_s \neq 0$, this differential equation is separable, and
can be solved analytically in terms of the {\em EllipticF} function
using Mathematica. Unlike in the impermeable
case~\cite{vee-raj-bir-pur2009}, the solution of~\eqref{eqn:curvODE} can
not be made periodic unless $\kappa(s) = \kappa_0$. Therefore, the
steady state shape of a semi-permeable vesicle in a quiescent flow is a
circle. By the inextensibility condition, the steady state vesicle has a
radius of $L/2\pi$, so its curvature and tension are
\begin{align}
  \kappa = \frac{2\pi}{L}, \: \Lambda = \frac{2\pi^2}{L^2}.
  \label{eqn:SSshape}
\end{align}
Note that in contrast to a impermeable vesicle, the final tension is
positive rather than negative.

In addition to computing the steady shape, we also compute the transient
area of the vesicle. Since the area of the vesicle is
\begin{align}
  A(t) = \frac{1}{2}\int_{\gamma} (\xx \cdot \nn)\, ds,
  \label{eqn:area}
\end{align}
its derivative includes terms due to the time derivative of $\xx \cdot
\nn$ and a term due to interface stretching~\cite{lai-tse-hua2008}, but
the stretching term is zero because of the local inextensibility
condition. Therefore,
\begin{align}
  \dot{A}(t) =
  \frac{1}{2} \int_{\gamma} (\dot{\xx} \cdot \nn)\, ds  + 
  \frac{1}{2} \int_{\gamma} (\xx \cdot \dot{\nn})\, ds.
\end{align}
Applying integration by parts to the second integral, it is equivalent
to the first.
%\begin{align}
%  \int_\gamma (\xx \cdot \dot{\nn})\, ds 
%    = \int_\gamma \left(\xx \cdot \dot{\xx}_s^\perp \right)
%    = -\int_\gamma \left(\xx^\perp \cdot \dot{\xx}_s\right)\, ds 
%    = \int_\gamma \left(\xx_s^\perp \cdot \dot{\xx}\right)\, ds 
%    = \int_\gamma (\dot{\xx} \cdot \nn)\, ds.
%\end{align}
Therefore,
\begin{align}
  \dot{A}(t) = \int_\gamma \left(\beta (\ff \cdot \nn)\nn 
    + \SS[\ff]\right) \cdot \nn\, ds 
  = \beta \int_\gamma (\ff \cdot \nn)\, ds,
\end{align}
where the last equality follows from the incompressibility of the
single-layer potential.  Using integration by parts and the
Frenet-Serret identities, we have
\begin{align}
  \dot{A}(t) &= \beta \int_\gamma \left( -\xx_{ssss} + 
    (\sigma \xx_s)_s \right) \cdot \nn \, ds \\
  &= -\beta \int_\gamma \left(-\xx_{sss} + \sigma \xx_s 
    \right) \cdot \kappa \xx_s \, ds \\
  &= -\beta \int_\gamma \left(\xx_{ss} \cdot 
    (\kappa \xx_s)_s + \kappa \sigma \right) \, ds \\
  &= -\beta \int_\gamma \left(\kappa (\xx_{ss} \cdot \xx_{ss}) + 
    \kappa_s (\xx_{ss} \cdot \xx_s) + \kappa \sigma \right) 
    \, ds \\
  &= -\beta \int_\gamma \left(\kappa^3 + \kappa \sigma \right) 
    \, ds = \beta \int_\gamma \left(
    \frac{\kappa^3}{2} - \kappa \Lambda \right) \, ds.
  \label{eqn:areaROC}
\end{align}
This is consistent with the steady state curvature and tension
in~\eqref{eqn:SSshape} since these constant values result in the
integrand in~\eqref{eqn:areaROC} being zero.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Discretization in Space}
Using an integral equation formulation, the unknowns are the vesicle's
position and tension that we track at a set of $N$ discretization
points. All derivatives are computed with Fourier differentiation, and
the single-layer potential operator is discretized with eighth-order
quadrature~\cite{alp1999}. For constricted flows, nearly-singular
integrals are accurately computed with an interpolation-based quadrature
rule~\cite{qua-bir2014}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Discretization in Time}
To eliminate stringent time step restrictions, we modify a time stepping
method that discretizes high-order derivatives
semi-implicitly~\cite{vee-gue-zor-bir2009}. Introducing notation for the
differential and projection operators
\begin{alignat}{3}
  &B[\xx]\ff = -\frac{d^4}{ds^4} \ff,  \qquad
  &&T[\xx]\sigma = (\sigma \xx_s)_s, \\
  &D[\xx]\ff = \xx_s \cdot \ff_s, 
  &&P[\xx]\ff = (\ff \cdot \nn) \nn,
\end{alignat}
the dynamics of the vesicle are governed by
\begin{align}
  \dot{\xx} = \uu_{\infty}(\xx) + 
  \beta P(B\xx + T\sigma) + \SS(B\xx + T\sigma), \quad
  D \dot{\xx} = 0.
\end{align}
Letting $B^N$ be the bending operator due to the vesicle configuration
at time $t^N$, and similar notation for the other operators, we use the
first-order semi-implicit time stepping method
\begin{alignat}{3}  
  \frac{\xx^{N+1} - \xx^N}{\Delta t} &= \uu_\infty(\xx^N) 
  &&+ \beta P^N\left(B^N\xx^{N+1} + T^N\sigma^{N+1}\right) \nonumber \\
  & &&+ \SS^N\left(B^N\xx^{N+1} + T^N\sigma^{N+1}\right), \\
  D^N\xx^{N+1} &= 1.
\end{alignat}

To achieve high-order accuracy over long time horizons, we modify an
adaptive spectral deferred correction time stepping
method~\cite{qua-bir2016}. Since the area of the vesicle is not
conserved, we only use the length of the vesicle to estimate the local
truncation error that determines if a time step is accepted or not and
the subsequent time step size.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Quiescent Flow} 
We consider a semi-permeable elliptical vesicle with initial reduced
area $\nu = 0.65$ and several different permeability rates in a
quiescent flow. The steady-state of an impermeable vesicle has constant
negative tension~\cite{kra-win-sei-lip1996}, but based
on~\eqref{eqn:SSshape}, a semi-permeable vesicle relaxes to a circle
with positive constant tension $\Lambda = 2\pi^2/L^2$. Recalling the
governing equation for the area in~\eqref{eqn:areaROC}, we expect both
the time for the vesicle to deviate from its initial area, and the time
to reach its steady state will scale with $\beta^{-1}$. In
Table~\ref{tbl:ellipseRelaxTimes}, we report the time,
$t_\mathrm{steady}$, required for the vesicle's area to be within 1.3\%
of its maximum area, and the time, $t_\mathrm{inflate}$, required for
the vesicle's area to exceed 1.3\% of its initial area. We see that both
times scale with $\beta^{-1}$. We also plot the area of a vesicle with
six different permeability rates in Figure~\ref{fig:ellipseArea}. The
area is computed directly from the vesicle shape (solid lines) and by
integrating~\eqref{eqn:areaROC} (marks) with a first-order
method. We also include black dashed lines indicating the times,
$t_{\mathrm{steady}}$, that the vesicle is near its steady state
circular shape.
\begin{table}[htp]
  \centering
  \caption{\label{tbl:ellipseRelaxTimes} The time for a semi-permeable
  vesicle to its steady state ($t_\mathrm{steady}$) and to begin
  inflating ($t_\mathrm{inflate}$).}
  \begin{tabular}{C{0.8cm}C{2cm}C{0.8cm}C{2cm}C{0.8cm}}
    $\beta$ & $t_\mathrm{steady}$ & ratio & $t_\mathrm{inflate}$ & ratio \\
    \midrule
    $10^{0}$  & $6.45 \times 10^{0}$ & ---  
              & $2.17 \times 10^{-2}$ & --- \\ 
    $10^{-1}$ & $2.00 \times 10^{1}$ & 3.10 & 
                $1.25 \times 10^{-1}$ & 5.76 \\
    $10^{-2}$ & $9.61 \times 10^{1}$ & 4.81 & 
                $1.59 \times 10^{0}$ & 12.72 \\
    $10^{-3}$ & $8.36 \times 10^{2}$ & 8.71 & 
                $2.19 \times 10^{1}$ & 13.77 \\
    $10^{-4}$ & $8.15 \times 10^{3}$ & 9.74 & 
                $2.14 \times 10^{2}$ & 9.77 \\
    $10^{-5}$ & $8.03 \times 10^{4}$ & 9.85 & 
                $2.20 \times 10^{3}$ & 10.14 \\
    \bottomrule
  \end{tabular}
\end{table}

\begin{figure}[htp]
  \centering
%  \ifTikz
%  \input{figures/ellipseArea.tikz}
%  \fi
  \includegraphics[width=0.8\linewidth]{figures/ellipseArea}
  \caption{\label{fig:ellipseArea} The area of an elliptical
  semi-permeable vesicle, with permeability rate $\beta$ in a quiescent
  flow. The solid lines compute the area using~\eqref{eqn:area}, while
  the marks are computed by integrating~\eqref{eqn:areaROC}. The dashed
  black lines are the times when the vesicle's area is within 1.3\% of
  its maximum value of $4\pi$.}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Shear Flow}
We consider a semi-permeable vesicle in the shear flow
$\uu_{\infty}(\xx) = \dot{\gamma} (y,0)$ with different permeability
rates $\beta$ and shear rates $\dot{\gamma}$. Depending on its reduced
area and viscosity contrast, two-dimensional vesicles either tank tread
or tumble~\cite{fin-lam-sei-gom2008,kra-win-sei-lip1996}. Since we are
considering semi-permeable vesicles, we assume there is no viscosity
contrast, and tumbling is not possible.

We initialize an elliptical vesicle with reduced area $\nu = 0.65$, and
with its long axis oriented in the $y$-direction. For all permeability
rates and shear rates, the semi-permeable vesicle tilts to an
inclination angle and undergoes tank treading behavior. However, because
of the semi-permeability, the initial and final reduced areas differ. In
Figure~\ref{fig:shearPhaseDiagram}, we plot the final shape of the
vesicle for various permeability rates, $\beta$, and shear rates
$\dot{\gamma}$. The shear rate has the largest effect on the final
reduced area.

\begin{figure}[htp]
  \centering
  %\ifTikz
  %\input{figures/shearPhaseDiagramRA2.tikz}
  %\fi
  \includegraphics[width=\linewidth]{figures/shearPhaseDiagramRA2.pdf}
  \caption{\label{fig:shearPhaseDiagram} The steady state shape of a
  semi-permeable vesicle with several different permeability rates
  $\beta$ and shear rates $\dot{\gamma}$.}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Poiseuille Flow}
We consider a semi-permeable vesicle in the Poiseuille flow $\uu =
U(1-(y/W)^2,0)$, where $U$ is the maximum imposed velocity. The length
of the vesicle is $L = 4\pi$ and its initial reduced area is $\nu =
0.65$. The width $W$ is chosen so that $W/R = 10$, where $R$ is the
radius of the vesicle. We investigate the effect of the maximum velocity
$U$ and the permeability rate $\beta$ on the steady state shape of a
vesicle initially placed on the center line $y=0$.
Figure~\ref{fig:parabolicPhaseDiagram} shows the steady state
configuration for a variety of maximum velocities and permeability
rates. Similar to the shear example, the maximum velocity strongly
affects the final shape while the permeability rate's effect is much
weaker. Instead, the permeability rate determines the time required for
the vesicle to reach a steady state. In
Table~\ref{tbl:parabolicRelaxTimes1}, we consider two maximum velocities
and report the time, $t_\mathrm{steady}$, required for the vesicle's
area to be within 1.3\% of its maximum area, and the time,
$t_\mathrm{inflate}$, required for the vesicle's area to exceed 1.3\% of
its initial area. We see that both times scale with $\beta^{-1}$.

\begin{figure}[htp]
%  \ifTikz
%  \input{figures/parabolicPhaseDiagram.tikz}
%  \fi
  \includegraphics[width=0.9\linewidth]{figures/parabolicPhaseDiagram.pdf}
  \caption{\label{fig:parabolicPhaseDiagram} The steady state shape of a
  semi-permeable vesicle with several different permeability rates
  $\beta$ and flow rates $U$. The numbers reported above each vesicle is
  its reduced area. At high flow rates, the vesicle self-intersects
  before reaching a steady state configuration.}
\end{figure}

\begin{table}[htp]
  \centering
  \caption{\label{tbl:parabolicRelaxTimes1} The time for a
  semi-permeable vesicle in a Poiseuille flow to reach its steady state
  ($t_\mathrm{steady}$) and to begin inflating ($t_\mathrm{inflate}$).
  The times are reported for the maximum velocities $0.1 \mu m/s$ and
  $10 \mu m/s$.}
  \begin{tabular}{C{0.8cm}C{2cm}C{0.8cm}C{2cm}C{0.8cm}}
    \multicolumn{5}{c}{$U = 0.1 \mu m/s$} \\
    $\beta$ & $t_\mathrm{steady}$ & ratio & $t_\mathrm{inflate}$ & ratio \\
    \midrule
    $10^{0}$  & $3.33 \times 10^{0}$ & ---  
              & $1.67 \times 10^{-2}$ & --- \\ 
    $10^{-1}$ & $1.37 \times 10^{1}$ & 4.12 & 
                $8.73 \times 10^{-1}$ & 5.23 \\
    $10^{-2}$ & $8.60 \times 10^{1}$ & 6.26 & 
                $1.11 \times 10^{0}$ & 12.74 \\
    $10^{-3}$ & $7.75 \times 10^{2}$ & 9.01 & 
                $1.63 \times 10^{1}$ & 14.70 \\
    $10^{-4}$ & $7.69 \times 10^{3}$ & 9.93 & 
                $1.74 \times 10^{2}$ & 10.63 \\
    $10^{-5}$ & $8.01 \times 10^{4}$ & 10.41 & 
                $1.66 \times 10^{3}$ & 9.58 \\
    \bottomrule
  \end{tabular}

  \begin{tabular}{C{0.8cm}C{2cm}C{0.8cm}C{2cm}C{0.8cm}}
    \multicolumn{5}{c}{$U = 10 \mu m/s$} \\
    $\beta$ & $t_\mathrm{steady}$ & ratio & $t_\mathrm{inflate}$ & ratio \\
    \midrule
    $10^{0}$  & $3.55 \times 10^{0}$ & ---  
              & $1.67 \times 10^{-2}$ & --- \\ 
    $10^{-1}$ & $1.87 \times 10^{1}$ & 5.27 & 
                $9.01 \times 10^{-2}$ & 5.40 \\
    $10^{-2}$ & $1.17 \times 10^{2}$ & 6.26 & 
                $1.52 \times 10^{0}$ & 16.87 \\
    $10^{-3}$ & $1.06 \times 10^{3}$ & 9.06 & 
                $2.07 \times 10^{1}$ & 13.62 \\
    $10^{-4}$ & $1.02 \times 10^{4}$ & 9.62 & 
                $2.25 \times 10^{2}$ & 10.87 \\
    $10^{-5}$ & $1.04 \times 10^{5}$ & 10.20 & 
                $2.25 \times 10^{3}$ & 10.00 \\
    \bottomrule
  \end{tabular}
\end{table}
